Modal type theory based on the intuitionistic epistemic logic

### 2018-01-13 article.tex
start to rewrite

  \begin{center}
  \begin{prooftree}
  \AxiomC{}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\Gamma , \alpha \vdash \alpha$}
  \end{prooftree}
  \end{center}

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  	\begin{prooftree}
  	\AxiomC{$\Gamma, \alpha \vdash \beta$}
  	\RightLabel{$\rightarrow_i$}
  	\UnaryInfC{$\Gamma \vdash \alpha \to \beta$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash \alpha$ }
  	\AxiomC{ $\Gamma \vdash \beta$ }
  	\RightLabel{$\land_i$}
  	\BinaryInfC{$\Gamma \vdash \alpha \land \beta$}
  	\end{prooftree}

  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash \alpha$}
    	\RightLabel{$\textbf{K}_I$}
    	\UnaryInfC{$\Gamma \vdash \textbf{K}\alpha$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    	\begin{prooftree}
  	\AxiomC{$\Gamma \vdash \alpha \to \beta$}
  	\AxiomC{$\Gamma \vdash \alpha$}
  	\RightLabel{$\rightarrow_i$}
  	\BinaryInfC{$\Gamma \vdash \beta$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash \alpha_1 \land \alpha_2$ }
  	\RightLabel{$\land_e, i \in \{ 1, 2 \}$}
  	\UnaryInfC{$\Gamma \vdash \alpha_i$}
  	\end{prooftree}

  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash \textbf{K}(\alpha \to \beta)$}
    	\AxiomC{$\Gamma \vdash \textbf{K}\alpha$}
    	\RightLabel{$\textbf{K}_{MP}$}
          \BinaryInfC{$\Gamma \vdash \textbf{K}\beta$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

  \begin{lemma}
    NIEL$^{-}_{\land, \to} \vdash \alpha \Rightarrow$ IEL$^{-}$ $\vdash \alpha$.
  \end{lemma}

@@

1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \alpha$, then $IEL^{-} \vdash \wedge \Gamma \rightarrow \textbf{K}\alpha$.

$\begin{array}{lll}
(1)& \wedge \Gamma \rightarrow \alpha & \text{assumption}\\
(2)& \alpha \rightarrow \textbf{K}\alpha &\text{co-reflection.}\\
(3)& (\wedge \Gamma \rightarrow \alpha) \rightarrow ((\alpha \rightarrow \textbf{K}\alpha) \rightarrow (\wedge \Gamma \rightarrow \textbf{K}\alpha))&\text{IPC theorem}\\
(4) &(\alpha \rightarrow \textbf{K}\alpha) \rightarrow (\wedge \Gamma \rightarrow \textbf{K}\alpha) &\text{1, 3, MP}\\
(5) &\wedge \Gamma \rightarrow \textbf{K}\alpha &\text{2,4, MP}\\
\end{array}$

@@

2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}\alpha$, then $IEL^{-} \vdash \wedge \Gamma \rightarrow \textbf{K}\beta$.

$\begin{array}{lll}
(1)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta) & \text{condition}\\
(2)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}\alpha & \text{condition}\\
(3)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha & \land_i \\
(4)& \wedge \Gamma \rightarrow \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha & \text{induction
hypothesis} \\
(5)& (\textbf{K}(\alpha \rightarrow \beta) \rightarrow (\textbf{K}\alpha \to \textbf{K}\beta)) \rightarrow
((\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta)& \text{IPC
theorem} \\
(6)&\textbf{K}(\alpha \rightarrow \beta) \rightarrow (\textbf{K}\alpha \to \textbf{K}\beta)&
\text{normality}\\
(7)&(\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta)& \text{5,6,
MP}\\
(8)&(\wedge \Gamma \rightarrow \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha) \rightarrow & \\
&\quad\rightarrow ((\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow
\textbf{K}\beta) \rightarrow
(\wedge \Gamma \rightarrow \textbf{K}\beta))& \text{IPC theorem} \\
(9) &(\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta) \rightarrow
(\wedge \Gamma \rightarrow \textbf{K}\beta)& \text{4, 8, MP} \\
(10) &\wedge \Gamma \rightarrow \textbf{K}\beta& \text{7, 9, MP}
\end{array}$

@@

  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{\textbf{K}}$ of terms is defined by the grammar:
  \begin{equation}
    \Lambda_{\textbf{K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{\textbf{K}}) \: | \:
    (\Lambda_{\textbf{K}}\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} , \Lambda_{\textbf{K}}) \: | \:
    (\pi_i \Lambda_{\textbf{K}}) \:
    | \: ({\bf pure \: } \: \Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} \star
    \Lambda_{\textbf{K}})
  \end{equation}
  where $i \in \{ 1, 2 \}$.
  \end{defin}

@@

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash f : \textbf{K}(\alpha \to \beta)$}
    \AxiomC{$\Gamma \vdash x : \textbf{K}\alpha$}
    \RightLabel{$\textbf{K}_{app}$}
    \BinaryInfC{$\Gamma \vdash f \star x : \textbf{K}\beta$}
  \end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

  $\textbf{K}_{app}$ is a typing rule for lifted application. This rule is similar to
  $\rightarrow_e$-rule, therefore we introduce the infix operator $\star$ for lifted application.

  \begin{con}

  \emph{$\star$ is a left-associative, so $M \star N \star P = (M \star N) \star P$,  unless otherwise specified.}

  \end{con}

@@

  \begin{prooftree}
  \AxiomC{$x : \alpha \vdash x : \alpha$}
  \RightLabel{$\textbf{K}_I$}
  \UnaryInfC{$x : \alpha \vdash {\bf pure \: } x : \textbf{K} \alpha$}
  \RightLabel{$\rightarrow_i$}
  \UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : \alpha \to \textbf{K}\alpha$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$f : \textbf{K}(\alpha \to \beta) \vdash f : \textbf{K}(\alpha \to \beta)$}
    \AxiomC{$x : \textbf{K}\alpha \vdash x : \textbf{K}\alpha$}
    \RightLabel{$\textbf{K}_{app}$}
    \BinaryInfC{$f : \textbf{K}(\alpha \to \beta), x : \textbf{K}\alpha \vdash f \star x :
    \textbf{K}\beta$}
    \RightLabel{$\rightarrow_i$}
    \UnaryInfC{$f : \textbf{K}(\alpha \to \beta) \vdash \lambda x. f \star x : \textbf{K}\alpha \to \textbf{K}\beta$}
    \RightLabel{$\rightarrow_i$}
    \UnaryInfC{$\vdash \lambda f. \lambda x. f \star x: \textbf{K}(\alpha \to \beta) \to
    \textbf{K}\alpha \to \textbf{K}\beta$}
  \end{prooftree}

>>>>

  \begin{center}
  \begin{prooftree}
  \AxiomC{}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\Gamma , \alpha \vdash A$}
  \end{prooftree}
  \end{center}

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  	\begin{prooftree}
  	\AxiomC{$\Gamma, A \vdash B$}
  	\RightLabel{$\rightarrow_i$}
  	\UnaryInfC{$\Gamma \vdash A \to B$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash A$ }
  	\AxiomC{ $\Gamma \vdash B$ }
  	\RightLabel{$\land_i$}
  	\BinaryInfC{$\Gamma \vdash A \land B$}
  	\end{prooftree}

  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{$\textbf{K}_I$}
    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    	\begin{prooftree}
  	\AxiomC{$\Gamma \vdash A \to B$}
  	\AxiomC{$\Gamma \vdash A$}
  	\RightLabel{$\rightarrow_i$}
  	\BinaryInfC{$\Gamma \vdash B$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash A_1 \land A_2$ }
  	\RightLabel{$\land_e, i \in \{ 1, 2 \}$}
  	\UnaryInfC{$\Gamma \vdash A_i$}
  	\end{prooftree}

  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
    \AxiomC{$\vec{A} \vdash B$}
    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

Where $\Gamma \vdash {\bf K} \vec{A}$ is a syntax sugar for $\Gamma \vdash {\bf K} A_1 ,\dots, \Gamma \vdash {\bf K} A_n$.

@@

  \begin{lemma}
    $\Gamma \vdash_{\text{NIEL}^{-}_{\land, \to}} A \Rightarrow$ IEL$^{-} \vdash \bigwedge \Gamma \rightarrow A$.
  \end{lemma}

@@

1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}A$.

$\begin{array}{lll}
(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
(2) & A \rightarrow \textbf{K}A &\text{co-reflection}\\
(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A))&\text{IPC theorem}\\
(4) & (A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A) &\text{from (1), (3) and MP}\\
(5) & \bigwedge \Gamma \rightarrow \textbf{K}A &\text{from (2), (4) and MP}\\
\end{array}$

@@

2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}B$.

$\begin{array}{lll}
(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL theorem} \\
(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (2), (3) and MP} \\
(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
\end{array}$

@@

  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{\textbf{K}}$ of terms is defined by the grammar:

$\begin{array}{lll}
& \Lambda_{\textbf{K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}}\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} , \Lambda_{\textbf{K}}) \: | \: (\pi_1 \Lambda_{\textbf{K}}) \: | \: (\pi_2 \Lambda_{\textbf{K}}) \: | & \\
& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{\textbf{K}}) \: | \: ({\bf let \:\: pure} \:\: \Lambda_{\textbf{K}} = \Lambda_{\textbf{K}} \:\: {\bf in} \:\: \Lambda_{\textbf{K}})
\end{array}$

@@

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash \vec{x} : {\bf K} \vec{A}$}
    \AxiomC{$\vec{y} : \vec{A} \vdash M : B$}
    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{y} = \vec{x} \:\: {\bf in} \:\: M : {\bf K} B$}
  \end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

@@

  \begin{prooftree}
  \AxiomC{$x : A \vdash x : A$}
  \RightLabel{$\textbf{K}_I$}
  \UnaryInfC{$x : A \vdash {\bf pure \: } x : \textbf{K} A$}
  \RightLabel{$\rightarrow_i$}
  \UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
    \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
    \AxiomC{$g : A \to B$}
    \AxiomC{$y : A$}
    \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
    \TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
    \UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
    \UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
  \end{prooftree}

\begin{prooftree}
  \AxiomC{$f : A \to B \vdash f : A \to B$}
  \UnaryInfC{$f : A \to B \vdash {\bf pure \:\:} f : {\bf K }(A \to B)$}
  \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
  \AxiomC{$g : A \to B$}
  \AxiomC{$y : A$}
  \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
  \TrinaryInfC{$f : A \to B, x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
  \UnaryInfC{$f : A \to B \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
  \UnaryInfC{$\lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : (A \to B) \to {\bf K} A \to {\bf K} B$}
\end{prooftree}
